---
title: "ggplot2"
author: "Jasmine"
date: '2019-04-29'
output: html_document
---

lesson modified from Data Carpentry [R-ecology-lesson](http://datacarpentry.org/R-ecology-lesson/04-visualization-ggplot2.html) 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# install.packages(c(ghibili, wesanderson, plotly, viridis))
```

Here is an easy visualization on how ggplot layers on and how the different commands using this [flipbook](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html#21)

# Download data
```{r}
library(tidyverse)

# Download the data
download.file(url="https://ndownloader.figshare.com/files/2292169",
              destfile = "portal_data_joined.csv")

surveys_complete <- read_csv("data/portal_data_joined.csv")

#view the first few rows
head(surveys_complete)
```

# make a new plot
a basic ggplot call is made up of:
1. `ggplot()`
2. `geom_[some kind of plot]()`
3. .. and more calls to make your plots more fancy or more complicated facets

where:
1. `ggplot()` --> holds your data that will be applied to all subsequent layers
2. `geom_[some kind of plot]()` --> determines the type of graph you have ie `geom_line` = line graph
3. and these calls are chained together by +

```{r}
test_plot <- ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) + 
  geom_hex()
```
_pro tip_: don't forget that the + needs to be at the end of each line

# Plotly
a way to look at data more easily in an interactive plot
- `ggplot` plot can be directly converted to a `plotly` plot
```{r}
library(plotly)

ggplotly()
ggplotly(test_plot)

last_plot()

ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) + 
  geom_point()
```


# a basic plot with coloured by species
- place arguments like `x`, `y` and `colour` in `aes()`
- arguments like `alpha` and `size` are applied to the general points so are outside `aes()`
- use `labs()` to make your axis and title look nice
```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length, colour = species)) + 
  geom_point(alpha = 0.5, size = 10) +
  geom_smooth(method = "lm", aes(group = 1)) +
  labs(title = "new title here \nhello",
       y = expression(paste("hindfoot length (cm^2)", mu, ")")),
       x = "weight (g)",
       colour = "rats") 
```

# Make you plots pretty with colours ~

## Ghibli

### View the palette
```{r}
library(ghibli)

ghibli_palettes

# display palettes w/ names
par(mfrow=c(9,3))
for(i in names(ghibli_palettes)) print(ghibli_palette(i))

pal <- ghibli_palette("KikiLight")

pal2 <- c(pal[3], pal[4])
```

### use the palette 
- ghibli has it's own function `scale_fill_ghibli_c` (continuous) or `_d` (discrete) 
- just add the name of the palette as an argument
```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) +
  geom_hex() +
  scale_fill_ghibli_c("PonyoLight")
```


## wesanderson
https://github.com/karthik/wesanderson
- this palette doesn't have it's own function yet 
```{r}
library(wesanderson)

# look at your palettes
wes_palettes

# assign the palette you want to use to something
pal <- wes_palette("Zissou1", 100, type = "continuous")

ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) +
  geom_hex() +
  scale_fill_gradientn(colours = pal)
```

## Viridis
https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html
- a nice package with colours that are colourblind friendly
- this is similar to how ghibli works but doesn't have a _c or _d function but instead uses either `discrete = TRUE` or `FALSE`
```{r}
yearly_counts <- surveys_complete %>%
                 count(year, species_id)

library(viridis)

ggplot(data = yearly_counts, mapping = aes(x = year, y = n, color = species_id)) +
    geom_line() 

ggplot(data = yearly_counts, mapping = aes(x = year, y = n, color = species_id)) +
    geom_line() +
  # add this line to use the viridis colour package
  # TRUE for discrete variable and FALSE for continuous variables
   scale_color_viridis(discrete = TRUE)

```

